---
title: "R Practice in Notebook"
output: Practice_notebook
---

This script is to practice R codes
Autor 
Rakib Al-Fahad 
Department of Electrical and Computer Engineering
The University of Memphis
TN-38152
ralfahad@memphis.edu

'Hi this is a test'

Set git
https://www.youtube.com/watch?v=hKoSJGWnFFA


Download R for MAC
https://cran.r-project.org/bin/macosx/

Download R for MAC
https://cran.r-project.org/bin/windows/


Download rstudio from website
https://www.rstudio.com/

To do:
1. Install R
2. Install RStudio


Let's begin with the Hellow world example

```{r}
print ('Hellow World')
```


Current directory and change the directory
```{r}
getwd()
setwd("/Users/rakib/LearnR/LearnRCode")

```

Write python code


```{python}
import sys
import pandas as pd 
a_py= pd.DataFrame(range(1,10))
print a_py
```

Writing Bash code
```{bash}
ls

```

Simple test code

```{r}
a<-5
b<-10
sum<-a+b
sum
print (sum)

```

You can use paste with print
```{r}
print(paste0("Sum is: ", sum))
#or cat
cat("Sum is: ", sum)
```

# R - Data Types
    Vectors
    Lists
    Matrices
    Arrays
    Factors
    Data Frames

```{r}
v <- TRUE 
print(class(v))

# as chr
v <- "TRUE"
print(class(v))

cat('\n')
cat('Test logic \n')
5==6
cat('\n')


v <- 23.5
print(class(v))

```



# Vectors
  When you want to create vector with more than one element, you should use c() function 
  which means to combine the elements into a vector.


```{r}
# Create a vector.
apple <- c('red','green',"yellow")
#apple <- c(2,5,6.3)
print(apple)

# Get the class of the vector.
print(class(apple))

```


# Lists
# A list is an R-object which can contain many different types of elements inside 
# it like vectors, functions and even another list inside it.

```{r}
# Create a list.
list1 <- list(c(2,5,3),21.3,sin)

cat('Print a element:\n')
list1[2]

cat('Print the list.\n')
print(list1)

```

# Create a matrix.
```{r}
cat('byrow: \n')
M = matrix( c('a','a','b','c','b','a'), nrow = 2, ncol = 3, byrow = TRUE)
print(M)

cat('\n')

cat('bycol: \n')
M = matrix( c('a','a','b','c','b','a'), nrow = 2, byrow = FALSE)
print(M)
```


# Create an array.
```{r}
a <- array(c('green','yellow'))
print (a)

cat('\n change dimention \n\n')

a <- array(c('green','yellow'),dim = c(3,3,2))
print(a)
```

# Create the data frame.

```{r}
BMI <- 	data.frame(
  gender = c("Male", "Male","Female"), 
  height = c(152, 171.5, 165), 
  weight = c(81,93, 78),
  Age = c(42,38,26)
)
print(BMI)

# dim(BMI)
# head(BMI, n=10)
# tail(BMI, n=2)

```

##################################### Finding Variables ######################


# Assignment using equal operator.


```{r}
var.1 = c(0,1,2,3)           
```

# Assignment using leftward operator.

```{r}
var.2 <- c("learn","R")   
```

# Assignment using rightward operator.   

```{r}
c(TRUE,1) -> var.3 
ls() # show all the variable in the enviornment
# List the variables starting with the pattern "var".
#print(ls(pattern = "var"))  
```




##################################### Delete Variables ######################
```{r}
rm(var.3)
print(var.3)
#rm(list=ls(all=TRUE)) # remove everything from a directory

```

##################################### R - Operators ######################


```{r}
# +,-,*,/
# %% Give the remainder of the first vector with the second
22%%7

# %//%	The result of division of first vector with second (quotient)
22%/%7

#^	The first vector raised to the exponent of second vector
a<-c(1,2,3)
a^a
a^2

#Relational Operators
# <, >, ==, <=, >=, !=
#Logical Operators
# &, | <-or, !<-Not, gives the opposite logical value
v <- c(3,0,TRUE,2+2i) 
print(!v)

#Assignment Operators
x<-5 # Called Left Assignment
x->z #Called Right Assignment
print(z)
```


##################################### MATH FUNCTIONS  ######################

```{r}
 
sqrt(x=100)
 
# Get the power you raise the base to get x
log(x=4, base=2)
 
# Euler's number 2.718 to the power of x
exp(x=2)
 
# Sum all vector values
sum(c(1,2,3))
 
# Find the mean (average)
randD1 = c(1,5,6,7,10,16)
mean(randD1)
 
# The median (Middle Number or avg of middle 2)
median(randD1)
 
# Minimum value
min(randD1)
 
# Maximum value
max(randD1)
 
# Min and max
range(randD1)
 
# Rounding
ceiling(4.5)
floor(4.5)
 
# Cumulatives
cat('\n')
print('Cumulatives')
cumsum(c(1,2,3))
cumprod(c(1,2,3))
cummax(c(7:9, 4:6, 1:3))
cummin(c(4:6, 1:3, 7:9))
 
# Generating Random samples
# Flipping a coin 10 times and weigh the probability
# of the next flip based on the previous
cat('\n')
print('Sample')
sample(0:10,5,replace=T)
sample(1:10,5,replace=F)


```




##################################### R-Decision making  ######################



```{r}
a<-10
if (a > 10){
  print("Greater")
}else if (a==10){
  print ('Equal')
}else{
  print ('Less')
}

#Sort form
x <- -5
y <- if(x > 0) 5 else 6
y

```
##################################### R - Loops ######################

Create a vector filled with random normal values


```{r}
u1 <- rnorm(30)
u1
#d <- density(u1)
#plot(d, main="Kernel Density of Miles Per Gallon")
#polygon(d, col="red", border="blue")

```

calculates the square of the first 10 elements of vector u1

```{r}

print("This loop calculates the square of the first 10 elements of vector u1")
# Initialize `usq`
usq <- 0

for(i in 1:10) {
  # i-th element of `u1` squared into `i`-th position of `usq`
  usq[i] <- u1[i]*u1[i]
  print(usq[i])
}

#print(i)
#print(u1)
```

print each element


```{r}
for(i in u1) {
  print(i)
}
```

##################################### Function Definition ######################

function_name <- function(arg_1, arg_2, ...) {
  Function body 
  }

Create a function that finds x to a user defined power


```{r}

power = function(exp=1){
  function(x){
    p<-x ^ exp
    return(p)
  }
}
 
cubed <-power()
cubed

cubed(2)
#cubed(1:5)


```


----- EXCEPTION HANDLING -----
Used to gracefully handle errors
I handle a division with string error



```{r}
divide = function(num1, num2){
  tryCatch(
    num1 / num2,
    error = function(e) {
      if(is.character(num1) || is.character(num2)){
        print("Can't Divide with Strings")
      }
  })
}
 
divide(10,'5')

```
 
 # ----- READING WRITING FILES -----
# Create a text file with headers fname lname sex
# and the data in a txt file Use `for missing values
# Save in the same directory as your R file
 
# Supply the file to read, whether the 1st line is 
# headers, what seperates the data, what is being used
# for missing data and false because you don't want to
# convert string vectors to factors
 
# myPeople is a data frame

```{r}
myPeople = read.table(file='test.txt',
                      header=T, sep=";",
                      na.strings="`",
                      stringsAsFactors=F)

myPeople

# Add another person
donnaRecord = data.frame(fname="Donna",
                         lname="Heyward",
                         sex="female")
myPeople = rbind(myPeople, donnaRecord)
myPeople

# Update a record
myPeople[7,2] = "Smith"
myPeople

# Update the file by supplying the data.frame,
# the file to write, seperator, na, whether to
# quote strings, whether to include row numbers

#write.table(x=myPeople, file='fil_name' # change the file name with extention
#            sep=" ", na="`",
#            quote=F, row.names=F)

```

Visualize the tables

```{r}
# Get 1st 3 records
head(myPeople, 3)
 
# Get remaining records
tail(myPeople, 3)
```

# ----- BASIC PLOTTING -----
# R provides great plotting tools
 
# Plotting x y coordinates from a matrix
# 1st 5 are x and 2nd 5 are y

```{r}

xy1 = matrix(data=c(1,2,3,4,5,
                    1,2,3,4,5), nrow=5, ncol=5)

xy1
#plot(xy1)

```

Draw a line

```{r}
x2 = c(1,2,3,4,5)
y2 = c(1,2,3,4,5)
#line
plot(x2, y2, type="l")

# Points and lines
plot(x2, y2, type="b")

# Points and lines with no space around points,
# labels, a blue line (Find more with colors())
plot(x2, y2, type="o", 
     main="My Plot", xlab="x axis", ylab="y axis",
     col="red")

# pch (1-25) defines different points
# lty (1-6) defines different lines
# xlim defines the max and min x plotting region
# ylim defines the max and min y plotting region
plot(x2, y2, type="b", pch=2, lty=2,
     xlim=c(-8,8), ylim=c(-8,8))
 
# Multiple plots
plot(x2, y2, type="b")
 
# Adds straight lines at 2 and 4 coordinates
abline(h=c(2,4), col="red",lty=2)
 
# Draw a 2 segmented lines with starting and ending x
# and y points
segments(x0=c(2,4), y0=c(2,2), x1=c(2,4), y1=c(4,4),
         col="red",lty=2)
 
# Draw an arrow
arrows(x0=1.5, y0=4.55, x1=2.7, y1=3.3, col="blue")
 
# Print Text
text(x=1.25, y=4.75, labels="Center")



```



PIE CHARTS 

```{r}
# Simple Pie Chart
#png(file="food_pref_bar_chart.png")
slices <- c(10, 12,4, 16, 8)
lbls <- c("US", "UK", "Australia", "Germany", "France")
pie(slices, labels = lbls, main="Pie Chart of Countries")
#dev.on()

```


# ----- REGRESSION ANALYSIS -----

```{r}

# Used to study a relationship between 2 separate 
# pieces of data (What is the relation between batting
# average and RBIS)
 
# Create relationship model between AVG and RBIs
mlbPlayers = read.table(file=file.choose(), #file='file_name'
                        header=T, sep=" ",
                        na.strings="`",
                        stringsAsFactors=F)
 
summary(mlbPlayers)


# Grab just RBIs and Avg for each player
# playerData is known as a data frame (Table of Data)
# We get the stats we want by passing that list in a vector
playerData = mlbPlayers[,c("RBI","AVG")]
playerData
relation = lm(playerData$RBI~playerData$AVG)
 
# Create file 
#png(file="RBI_AVG_Regression.png")
 
# Plot the chart
plot(playerData$AVG, playerData$RBI, 
     main="AVG & RBI Regression", 
     abline(lm(playerData$RBI~playerData$AVG)),
     xlab="AVG", ylab="RBIs")

plot(relation) 
# Save chart
#dev.off()
```


# ----- MULTIPLE REGRESSION -----
# Used to study the impact on one variable from numerous 
# others
# Estimate RBIs based on other player stats

```{r}
 
playerData2 = mlbPlayers[,c("RBI","AVG","HR","OBP",
                            "SLG","OPS")]
 
# Create the relationship model
relation2 = lm(playerData2$RBI ~ playerData2$AVG + 
                 playerData2$HR + playerData2$OBP +
                 playerData2$SLG + playerData2$OPS)
 
sprintf("Intercept : %f1.4", coef(relation2)[1])
 
# How stats effect RBIs
sprintf("AVG : %f1.4", coef(relation2)[2])
sprintf("HR : %f1.4", coef(relation2)[3])
sprintf("OBP : %f1.4", coef(relation2)[4])
sprintf("SLG : %f1.4", coef(relation2)[5])
sprintf("OPS : %f1.4", coef(relation2)[6])
 
# Calculate expected RBIs based on stats
# Evan Longoria
# RBIs   AVG   HR   OBP   SLG   OPS
# 86     .261  20  .313  .424  .737
RBIGuess = -5.05 + (372.96 * .261) + (2.56 * 20) +
  (-5.41 * .313) + (-167.37 * .424)
RBIGuess
```

Bonus

Correlation plot


```{r}
library(corrplot)
head(mtcars)
M<-cor(mtcars)
M
corrplot(M, method="circle")

```


Apply functions

The apply() functions form the basis of more complex combinations and helps to perform operations with very few lines of code. More specifically, the family is made up of the 

apply(), lapply() , sapply(), vapply(), mapply(), rapply(), and tapply() functions.

link: https://www.datacamp.com/community/tutorials/r-tutorial-apply-family

```{r}
# Construct a 5x6 matrix
X <-matrix(rnorm(30), nrow=5, ncol=6)
X

# Sum the values of each column with `apply()`
A1<-apply(X, c(2), sum)
A1
# varify 
sum(X[,1])

# lapply
x <- list(a = 1:10)
# compute the list mean for each list element
B1<-lapply(x, sum)
print ('Lapply')
B1

```


